---
title: "InitialSamplingVisits_Code"
author: "Tim L"
date: "7/8/2023"
output: git_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
###Load Packages
library(dplyr)
library(ggplot2)
library(tidyverse)
library(lme4)
library(lmerTest)
library(rstatix)
library(emmeans)
library(car)
library(wesanderson)
library(forcats)
library(brglm2)
library(effectsize)
library(lsr)
```

###Read Data
```{r}
##CSV file with raw data
rawdata <- read.csv("InitialSamplingVisits_Data.csv", na.strings = c("", "NA"))
```


###Transform and summarize duplicate platings
```{r}
#set detection limit, then log transform, assign TNTC and atypical to actual numbers
pc_log <- rawdata %>% 
  mutate(Concentration = replace(Concentration, Concentration==0, 5)) %>% 
  mutate(Concentration = replace(Concentration, TNTC_Atyp == "TNTC", 20000)) %>% 
  mutate(Concentration = replace(Concentration, TNTC_Atyp == "atypical", 5)) %>% 
  mutate(Concentration = log10(Concentration)) 

#mean of spc and cvta duplicates, assign GN spoilage
pc_summary <- pc_log %>% 
  group_by(Plant, Trial, MilkType, BME, BME_Rep, Day, PlateType) %>% 
  drop_na(PlateType) %>% 
  summarise(Concentration = mean(Concentration)) %>% 
  spread(PlateType, Concentration) %>% 
  mutate(GNspoilage = if_else(CVTA > log10(5) & SPC >= log10(20000), "Y", "N")) %>%
  drop_na(GNspoilage) %>% 
  drop_na(CVTA) %>% 
  drop_na(SPC)


#Gram negative contamination assigned to plates with detectable growth on CVTA
pc_log_summary_GN <- pc_log %>% 
  group_by(Plant, Trial, MilkType, BME, BME_Rep, Day, PlateType) %>% 
  drop_na(PlateType) %>% 
  summarise(Concentration = mean(Concentration)) %>% 
  spread(PlateType, Concentration) %>% 
  mutate(GramNeg = if_else(CVTA > log10(5), "Y", "N")) %>%
  drop_na(GramNeg) %>% 
  drop_na(CVTA) %>% 
  drop_na(SPC)
```

###Descriptive Stats
```{r}
#mean spc and cvta grouped by plant, trial, and day
pc_PlantTrialDay <- pc_summary %>% 
  ungroup() %>% 
  group_by(Plant, Trial, Day) %>% 
  summarise_at(c("SPC", "CVTA"), mean, na.rm = TRUE)

#mean spc and cvta grouped by trial and day
pc_TrialDay <- pc_summary %>% 
  ungroup() %>% 
  group_by(Trial, Day) %>% 
  summarise_at(c("SPC", "CVTA"), mean, na.rm = TRUE)

#mean spc and cvta grouped by day
pc_Day <- pc_summary %>% 
  ungroup() %>% 
  group_by(Day) %>% 
  summarise_at(c("SPC", "CVTA"), mean, na.rm = TRUE)

#mean spc grouped by day and gram-negative contamination
pc_DayGN <- pc_log_summary_GN %>% 
  ungroup() %>% 
  group_by(Day, GramNeg) %>% 
  summarise_at(c("SPC"), mean, na.rm = TRUE)

#stdv of spc grouped by day and gram-negative contamination
pc_DayGN_stdv <- pc_log_summary_GN %>% 
  ungroup() %>% 
  group_by(Day, GramNeg) %>% 
  summarise_at(c("SPC"), sd, na.rm = TRUE)
```

###Day 7 and 14 data subset and individual trial plate counts summaries
```{r}
#subset day 7 and 14 plate counts
pc_summary_d7d14 <- pc_summary %>% 
  filter(Day != 0) %>% 
  mutate(Day = factor(Day))

#subset trial 1 plate counts
pc_summary_t1 <- pc_summary %>% 
  filter(Trial == 1)

#Trial 1 mean plate counts by day
pc_Day_t1 <- pc_summary_t1 %>% 
  ungroup() %>% 
  group_by(Day) %>% 
  summarise_at(c("CVTA", "SPC"), mean, na.rm = TRUE)

#subset trial 2 plate counts
pc_summary_t2 <- pc_summary %>% 
  filter(Trial == 2)

#Trial 2 mean plate counts by day
pc_Day_t2 <- pc_summary %>% 
  ungroup() %>% 
  group_by(Day) %>% 
  summarise_at(c("CVTA", "SPC"), mean, na.rm = TRUE)
```

###Sensory data
```{r}
#filter out sensory data from raw data
sensory_summary <- rawdata %>% 
  drop_na(OVERALL.FLAVOR.SCORE) %>% 
  mutate(flavorscore = OVERALL.FLAVOR.SCORE) %>% 
  group_by(Plant, Trial, MilkType, BME, BME_Rep, Day) %>% 
  summarise(flavorscore = mean(flavorscore)) %>% 
  mutate(Day = as.factor(Day))

#summarize sensory by plant, trial, day
sensory_PlantTrialDay <- sensory_summary %>% 
  group_by(Plant, Trial, Day) %>% 
  summarise_at(c("flavorscore"), mean, na.rm = TRUE)

#summarize by trial and day
sensory_TrialDay <- sensory_summary %>% 
  group_by(Trial, Day) %>% 
  summarise_at(c("flavorscore"), mean, na.rm = TRUE)

#summarize by day
sensory_Day <- sensory_summary %>% 
  group_by(Day) %>% 
  summarise_at(c("flavorscore"), mean, na.rm = TRUE)

#mean flavor scores for trial 1
sensory_t1 <- sensory_summary %>% 
  filter(Trial == 1) %>% 
  ungroup() %>% 
  group_by(Day) %>% 
  summarise(flavorscore = mean(flavorscore))

#mean flavor scores for trial 2
sensory_t2 <- sensory_summary %>% 
  filter(Trial == 2) %>% 
  ungroup() %>% 
  group_by(Day) %>% 
  summarise(flavorscore = mean(flavorscore))
```


###combine plate count and sensory summary data
```{r}
#combine plate count and sensory data
pc_sensory_summary <- merge(pc_log_summary_GN, sensory_summary)

#mean sensory by day and gn contamination
sensory_DayGN <- pc_sensory_summary %>% 
  ungroup() %>% 
  group_by(Day, GramNeg) %>% 
  summarise_at(c("flavorscore"), mean, na.rm = TRUE)

#stdv sensory by day and gn contamination
sensory_DayGN_stdv <- pc_sensory_summary %>% 
  ungroup() %>% 
  group_by(Day, GramNeg) %>% 
  summarise_at(c("flavorscore"), sd, na.rm = TRUE)
  
#subset for days 7 and 14
pc_sensory_d7d14 <- pc_sensory_summary %>% 
  filter(Day == 7 | Day == 14) %>% 
  mutate(Day = as.factor(Day)) %>% 
  mutate(Trial = as.factor(Trial))
```


###logistic regression analysis for GN spoilage
```{r}
#convert predictor variables to factors
pc_summary_d7d14$GNspoilage <- as.factor(pc_summary_d7d14$GNspoilage)
pc_summary_d7d14$Plant <- as.factor(pc_summary_d7d14$Plant)
pc_summary_d7d14$BME_Rep <- as.factor(pc_summary_d7d14$BME_Rep)
pc_summary_d7d14$Trial <- as.factor(pc_summary_d7d14$Trial)

#create logistic (binomial) regression model for trial 1
model_T1 <- glm(GNspoilage ~ Plant + MilkType + BME, data = subset(pc_summary_d7d14, Trial == 1), family = binomial, method = brglmFit)

#summarize trial 1 model
summary(model_T1)

#trial 1 pairwise comparisons
emmeans(model_T1, pairwise~MilkType, type = "response")

emmeans(model_T1, pairwise~Plant, type = "response")

emmeans(model_T1, pairwise~BME, type = "response")

#create model for trial 2
model_T2 <- glm(GNspoilage ~ Plant + MilkType + BME + BME_Rep, data = subset(pc_summary_d7d14, Trial == 2), family = binomial, method = brglmFit)

#summarise trial 2 model
summary(model_T2)

#pairwise comparisons for trial 2
emmeans(model_T2, pairwise~MilkType, type = "response")

emmeans(model_T2, pairwise~Plant, type = "response")

emmeans(model_T2, pairwise~BME, type = "response")

emmeans(model_T2, pairwise~BME_Rep, type = "response")

emmeans(model_T2, pairwise~Plant|MilkType, type = "response")

#create model for trial 1 and 2 combined
model_T1T2 <- glm(GNspoilage ~ Plant + MilkType + BME + Trial, data = pc_summary_d7d14, family = binomial, method = brglmFit)

#summarise combined trial model
summary(model_T1T2)

#pairwise comparisons for trials 1 and 2 combined
emmeans(model_T1T2, pairwise~MilkType, type = "response")

emmeans(model_T1T2, pairwise~Plant, type = "response")

emmeans(model_T1T2, pairwise~BME, type = "response")

emmeans(model_T1T2, pairwise~Trial, type = "response")
```


###ANOVAs for sensory and plate counts
```{r}
#SPC anova for days comparing samples with and without GN spoilage on 7 and 14
#don't include 0 because no samples had GN spoilage on day 0
aov_SPC_d7d14 <- glm(SPC ~ Day + GramNeg + (Day*GramNeg) + MilkType, data = pc_summary_d7d14_GN)

#summaries
summary(aov_SPC_d7d14)

aov(aov_SPC_d7d14)

#pairwise comparisons
emmeans(aov_SPC_d7d14, pairwise ~ GramNeg)
emmeans(aov_SPC_d7d14, pairwise ~ Day)
emmeans(aov_SPC_d7d14, pairwise ~ MilkType)
emmeans(aov_SPC_d7d14, pairwise ~ GramNeg | Day)

#effect size
effectsize(aov(aov_SPC_d7d14))

#day 7 and 14 sensory anova comparing GN spoilage, day, and milk type
aov_sensory <- glm(flavorscore ~ GramNeg + Day + MilkType, data = pc_sensory_d7d14)

#summaries
summary(aov_sensory)

aov(aov_sensory)

#effect size
effectsize(aov(aov_sensory))

#pairwise comparisons
emmeans(aov_sensory, pairwise ~ GramNeg | Day)
emmeans(aov_sensory, pairwise ~ Day)
emmeans(aov_sensory, pairwise ~ MilkType)
```

###Grouped GN Frequencies
```{r}
#GN spoilage frequency by plant, trial, day
GNfreq_PlantTrialDay <- pc_summary %>%
  group_by(Plant, Trial, Day, GNspoilage) %>% 
  summarise(n = length(GNspoilage)) %>%
  mutate(n = replace_na(n, 0)) %>% 
  mutate(freq = 100*(n / sum(n)))

#GN spoilage frequency by trial, milk type, day
GNfreq_MilkTypeTrialDay <- pc_summary %>%
  group_by(MilkType, Trial, Day, GNspoilage) %>% 
  summarise(n = length(GNspoilage)) %>%
  mutate(n = replace_na(n, 0)) %>% 
  mutate(freq = 100*(n / sum(n)))

#GN spoilage frequency by BME, trial, milk type, day
GNfreq_TimePointMilkTypeTrialDay <- pc_summary %>%
  group_by(BME, Trial, Day, GNspoilage) %>% 
  summarise(n = length(GNspoilage)) %>%
  mutate(n = replace_na(n, 0)) %>% 
  mutate(freq = 100*(n / sum(n)))

#GN spoilage frequency by lane, trial, milk type, day
GNfreq_LaneMilkTypeTrialDay <- pc_summary %>%
  group_by(BME_Rep, Trial, Day, GNspoilage) %>% 
  summarise(n = length(GNspoilage)) %>%
  mutate(n = replace_na(n, 0)) %>% 
  mutate(freq = 100*(n / sum(n)))


#GN spoilage by trial and day
GNfreq_TrialDay <- pc_summary %>%
  group_by(Trial, Day, GNspoilage) %>% 
  summarise(n = length(GNspoilage)) %>%
  mutate(n = replace_na(n, 0)) %>% 
  mutate(freq = 100*(n / sum(n)))

#GN spoilage by day
GNfreq_Day <- pc_summary %>%
  group_by(Day, GNspoilage) %>% 
  summarise(n = length(GNspoilage)) %>%
  mutate(n = replace_na(n, 0)) %>% 
  mutate(freq = 100*(n / sum(n)))
```

###Supplemental Figure 2, Day 14 GN frequencies
```{r}
##BY FACILITY

#GN spoilage by plant day 14
GNfreq_PlantD14 <- pc_summary %>%
  filter(Day == 14) %>% 
  group_by(Plant, Day, GNspoilage) %>% 
  summarise(n = length(GNspoilage)) %>%
  mutate(n = replace_na(n, 0)) %>% 
  mutate(freq = 100*(n / sum(n))) %>% 
  filter(GNspoilage == "Y")

#plot
ggplot(GNfreq_PlantD14, aes(Plant)) +
  geom_col(aes(y = freq)) +
  ylab("Gram-negative spoilage frequency (%)") +
  ylim(0,100) +
  xlab("Facility") +
  labs(fill = "Plant", title = "(a)") +
  theme(plot.title = element_text(size=14, face="bold"))
          
#ggsave("FacilityGnfreq063023.jpeg", device = "jpg", type = "cairo", dpi = 600, width = 3.5, height = 3, units = "in")

##BY MILK TYPE

GNfreq_MilkTypeD14 <- pc_summary %>%
  filter(Day == 14) %>% 
  group_by(Plant, MilkType, GNspoilage) %>% 
  summarise(n = length(GNspoilage)) %>% 
  mutate(n = replace_na(n, 0)) %>% 
  mutate(freq = 100*(n / sum(n))) %>% 
  filter(GNspoilage == "Y")

ggplot(GNfreq_MilkTypeD14, aes(MilkType)) +
  geom_col(aes(y = freq)) +
  facet_grid(~Plant) +
  ylab("Gram-negative spoilage frequency (%)") +
  ylim(0,100) +
  xlab("Milk Type") +
  labs( title = "(b)") +
  scale_x_discrete(limits = rev) +
  theme(plot.title = element_text(size=14, face="bold"),   axis.text.x = element_text(angle = 60, hjust = 1))

#ggsave("MilkTypeGNfreq063023.jpeg", device = "jpg", type = "cairo", dpi = 600, width = 3.5, height = 3, units = "in")


##BY SAMPLING TIME POINT (Trial 1 only)

#GN spoilage by sampling time point and plant day 14
GNFreq_PlantTimepointD14T1 <-  pc_summary %>% 
  filter(Day == 14) %>% 
  filter(Trial == 1) %>% 
  group_by(Plant, BME, GNspoilage) %>% 
  summarise(n = length(GNspoilage)) %>%
  ungroup() %>% 
  complete(Plant, BME, GNspoilage) %>% 
  mutate(n = replace_na(n, 0)) %>% 
  group_by(Plant, BME) %>% 
  mutate(freq = 100*(n / sum(n))) %>% 
  filter(GNspoilage == "Y")

#order BME for plot
GNFreq_PlantTimepointD14T1$BME <- factor(GNFreq_PlantTimepointD14T1$BME, levels = c("B","M", "E"))
  
  
#plot Trial 1 (BME), separated by plant
ggplot(GNFreq_PlantTimepointD14T1, aes(BME)) +
  geom_col(aes(y = freq)) +
  facet_grid(~Plant) +
  ylab("Gram-negative spoilage frequency (%)") +
  ylim(0,100) +
  xlab("Sampling Time Point") +
  labs(fill = "Plant", title = "(c)") +
  facet_grid(~Plant) +
  theme(plot.title = element_text(size=14, face="bold")) + 
  scale_fill_manual(values = wes_palette("Darjeeling1")) +
  guides(fill=guide_legend(title="Facility"))

#ggsave("TimepointGNfreq063023.jpeg", device = "jpg", type = "cairo", dpi = 600, height = 3, width = 3.5, units = "in")


##BY LANE FOR TRIAL 2
GNFreq_PlantLaneD14T2 <- pc_summary %>%
  filter(Trial == 2) %>% 
  filter(Day == 14) %>% 
  group_by(Plant, BME_Rep, GNspoilage) %>% 
  summarise(n = length(GNspoilage)) %>% 
  mutate(n = replace_na(n, 0)) %>% 
  mutate(freq = 100*(n / sum(n))) %>% 
  filter(GNspoilage == "Y")


ggplot(GNFreq_PlantLaneD14T2, aes(BME_Rep)) +
  geom_col(aes(y = freq)) +
  facet_grid(~Plant) +
  ylab("Gram-negative spoilage frequency (%)") +
  ylim(0,100) +
  xlab("Lane Number") +
  labs(title = "(d)") +
  #scale_x_discrete(limits = rev) +
  theme(plot.title = element_text(size=14, face="bold")) + 
  scale_fill_manual(values = wes_palette("Darjeeling1"))


#ggsave("LaneGNfreq063023.jpeg", device = "jpg", type = "cairo", dpi = 600, height = 3, width = 3.5, units = "in")
```


###PPC vs Downtime
```{r}
#Filter out day 14 GN spoilage
GNfreq_PlantTrialD14 <- pc_summary %>%
  filter(Day == 14) %>% 
  group_by(Plant, Trial, GNspoilage) %>% 
  summarise(n = length(GNspoilage)) %>% 
  ungroup() %>% 
  group_by(Plant, Trial) %>% 
  mutate(freq = 100*(n / sum(n))) %>% 
  rename("n_samples" = "n") %>% 
  filter(GNspoilage == "Y") %>% 
  mutate(Plant = as.factor(Plant))

#Sum down time by plant and trial
downtime <- rawdata %>% 
  drop_na(DownTime) %>% 
  group_by(Plant, Trial) %>% 
  summarise(n_downtimes = sum(DownTime_n)) %>% 
  mutate(Plant = as.factor(Plant))

#Sum down time by plant, trial, General Down Time (Filler, non-Filler)
downtimegen <- rawdata %>% 
  drop_na(DownTime) %>% 
  group_by(Plant, Trial, GenDownTime) %>% 
  summarise(n_downtimes = sum(DownTime_n))

# Join GNspoilage and Downtime Tables
downtimeGN <- merge(GNfreq_PlantTrialD14, downtime)

downtimegenGN <- merge(GNfreq_PlantTrialD14, downtimegen) 

#Plot GNspoilage vs Downtime total
ggplot(downtimeGN, aes(freq)) +
  geom_point(aes(y = n_downtimes, shape = factor(Trial), colour = factor(Plant), size = 1)) +
  labs(title = "(a)", colour = "Plant", shape = "Trial") +
  ylab("Number of Occurences") +
  ylim(0,30) +
  xlim(0, 100) +
  xlab("Gram-negative Spoilage Frequency (%)") +
  theme(plot.title = element_text(size=14, face="bold")) +
  guides(colour = guide_legend("Facility"), shape = guide_legend("Sampling Visit"), size = FALSE) 

#ggsave("DownTimeAll063023.jpeg", device = "jpg", type = "cairo", dpi = 600)

#Plot GNspoilage vs General Down Time
ggplot(downtimegenGN, aes(freq)) +
  geom_point(aes(y = n_downtimes, shape = factor(Trial), colour = factor(Plant), size = 4)) +
  facet_grid(~GenDownTime) +
  labs(title = "(b)") +
  ylab("Number of Occurences") +
  ylim(0,30) +
  xlim(0, 100) +
  xlab("Gram-negative Spoilage Frequency (%)") +
  theme(plot.title = element_text(size=14, face="bold")) +
  guides(colour = guide_legend("Facility"), shape = guide_legend("Sampling Visit"), size = FALSE) 

#ggsave("DownTimeGen063023.jpeg", device = "jpg", type = "cairo", dpi = 600)
``` 